### 2014-04-17 brains-protocol.tex
Rework introduction of long-term evolution, needs comments on results

To settle this question we bin the quote occurrences into fixed \emph{time bags} spanning $\Delta t$ days (2 days in the implementation), each one representing a unit of time evolution
We further assume that when a quotation $q$ appears in time bag $i$, if it is not original (i.e. if not stemming from a source external to the dataset, e.g. initiating a new set of quotations), then it is based solely on the most frequent quotation $q_{max}$ in the preceding time bag ($i - 1$).
If $q$ differs from $q_{max}$ by only one word it is counted as a substitution from $q_{max}$ to~$q$.
In any other case, i.e. if $q$ and $q_{max}$ are the same or if $q$ and $q_{max}$ are different in other ways, the occurrence of $q$ is not considered to be an instance of substitution and is discarded.

@@

\section{Long-term evolution of quotation families}

@@

A discrete time is best suited to observe this: we first slice each quotation family into a number of time bags, each bag $\mathcal{B}_n$ containing all the quotations that appear in its corresponding time interval $\left[ t_n, t_{n + 1} = t_n + \Delta t \right[$ (where $\Delta t$ is a number of days, 2 in practice).
Each quote in a time bag is counted as many times as it occurs in the corresponding time interval.

@@

The raw distributions built with this computation are stationary.
\rk{figure/stats à l'appui ?
bouger ça dans les résultats ?
ou le garder comme ça, mentionné en passant pour justifier la suite ?}
That is, the substitutions on quotations over time have no global effect on quotation families, either because external quotations are continuously fed into the family and compensate for the effect of substitutions, or because substitutions operate on a marginal portion of the quotation families, or both.

\bigskip
To narrow this view to the specific effects of cognitive bias we consider substitution \emph{chains}: using the above-mentioned discrete-time substitution detection model, we filter time bags to include only new quotes produced by substitutions themselves based on quotes produced by substitutions, and so on recursively.
The first time bag is therefore untouched, the second contains the quotations produced by substitutions from the first, the third by substitutions from the second, and so on. \new{As a result, the size of observed words/quotes drastically decreases across time, yet it assuredly focuses on successive mutations.} \cam{ajouter graphique pédagogique?}
We thus observe only the repeatedly-affected portion of the quotation families, and obtain a simple view of the longitudinal effect of cognitive bias, \hbox{i.e.} how quote features are evolving in the long term and, perhaps, converging towards specific attractors.

@@

\section{Substitution process}

>>>>

To settle this question we bin the quote occurrences into fixed \emph{time bags} spanning $\Delta t$ days (2 days in the implementation), each one representing a unit of time evolution
We further assume that when a quotation $q$ appears in time bag $n$, if it is not original (i.e. if not stemming from a source external to the dataset, e.g. initiating a new set of quotations), then it is based solely on the most frequent quotation $q_{max}$ in the preceding time bag ($n - 1$).
If $q$ differs from $q_{max}$ by only one word it is counted as a substitution from $q_{max}$ to~$q$.
In any other case, i.e. if $q$ and $q_{max}$ are the same or if $q$ and $q_{max}$ are different in other ways, the occurrence of $q$ is not considered to be an instance of substitution and is discarded.

@@

\section{Macroscopic evolution of quotation families}

@@

To do so we use a variation of the model presented in the previous section: quote occurrences are binned into 2-day time bags, substitutions can stem from any quote in the time bag preceding the current one, but destination quotes must be absent from that preceding time bag. The last two points constitute a variation from the ``majority'' rule which let us detect substitution chains later on.

@@

The raw distributions built with this computation are stationary.
That is, the substitutions on quotations over time have no global effect on quotation families, either because external quotations are continuously fed into the family and compensate for the effect of substitutions, or because substitutions operate on a marginal portion of the quotation families, or both.

\bigskip
To narrow this view to the specific effects of cognitive bias we consider substitution \emph{chains}: using the above-mentioned substitution detection model, we filter time bags to include only new quotes produced by substitutions themselves based on quotes produced by substitutions, and so on recursively.
The first time bag is therefore untouched, the second contains the quotations produced by substitutions from the first, the third by substitutions from the second, and so on.
As a result, the size of observed words drastically decreases across time, yet it unambiguously focuses on successive mutations.
\rk{add pedagogical figure}
We thus observe only the repeatedly-affected portion of the quotation families, and obtain a simple view of the longitudinal effect of cognitive bias, \hbox{i.e.} how quote features are evolving in the long term and, perhaps, converging towards specific attractors.

\rk{comment results, transition to micro process}

@@

\section{Microscopic evolution: the substitution process}
